---
title: "L07 - Solutions"
author: "MAJ Kyle McDermott"
date: "2023-01-25"
output:
  pdf_document: default
  html_document: default

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE}
library(tidyverse)
library(janitor)
```

## Lesson 07 - Generalization (2.1) - SOLUTIONS

**Question 1**


b: What is the expected standard deviation of your distribution?
```{r 1b}
pi <- 0.40
n <- 25

sd <- sqrt((pi * (1 - pi))/n)
sd
```

**Question 3**


a: What is the expected standard deviation of your distribution?
```{r 3a}
pi <- 0.20
n <- 1414

sd <- sqrt((pi * (1 - pi))/n)
sd
```


b: How many standard deviations below the mean of the distribution described in part (a) is the sample proportion from the GSS?
```{r 3b}
phat <- 271/n
phat

z <- (phat - pi)/sd
z
```
d: Calculate a p-value for an alternative hypothesis that the true proportion of adults that have been told by a health professional that they had depression is not 20$\%$
```{r}
pvalue <- 2*(1-pnorm(abs(z)))

```

**Question 4**

Load in csv
```{r, message=FALSE}
Results <- read_csv("Starburst.csv")
```

Look data
```{r}
head(Results)
```

Plot data
```{r}
Results %>% 
  ggplot(aes(x = Worst)) +
  geom_bar()
```
Calculate proporations
```{r}
Results %>% 
  tabyl(Worst) %>% 
  adorn_totals()
```
```{r}
n <- 56
phat <- 17/n
pi <- (1/3)

sd <- sqrt((pi * (1-pi))/n)

z <- (phat - pi)/sd

pvalue <- 1-pnorm(z)

phat
sd
z
pvalue
```

**Question 5**
```{r}
n <- 140
phat <- 73/n
pi <- 0.40

sd <- sqrt((pi * (1 - pi))/n)
z <- (phat - pi)/sd
pvalue <- 2*(1-pnorm(abs(z)))

z
pvalue

```

